\section{Project Description}

The Department of Maintenance and Operations (Beheer en Onderhoud - B\&O) of the province of Gelderland manages a wide range of civil engineering assets, referred to as engineering structures (“Kunstwerken”), such as bridges, viaducts, and culverts. 
For these assets, original construction drawings and related technical documentation are frequently required, for example when performing structural recalculations, maintenance planning, or safety assessments. 
Currently, asset specialists spend a significant amount of time searching for relevant archival information. 
The documentation is stored across multiple archives on an internal network drive, each using different naming conventions, classifications, and terminology. 
There is no effective or universal search functionality available.
As a result, for each individual asset, specialists must search multiple archives using different search terms, without certainty about which archive contains the required information, or whether the information exists at all. 
This makes the retrieval of documentation a time-consuming and inefficient “treasure hunt.”  
Recent advances in natural language processing (NLP) and information retrieval (IR) techniques offer promising advantages for improving the search and retrieval of documents in such complex archival systems.
Additionally, techniques for extracting and structuring information from unstructured data, such as optical character recognition (OCR) for scanned historical documents, can help in making the content of these archives more accessible and searchable.

Here, we outline the plan for a project that strives to combine these techniques to implement a proof of concept for a ML-aided document retrieval system, to support the asset specialists from the province of Gelderland in their work.


\subsection{Background of the Project}

To begin with, we briefly review the established techniques in the fields of NLP, IR, and OCR which are possible candidates for our project, as well as the current state-of-the-art implementations of these.

OCR techniques have been established for decades and have seen significant improvements in accuracy and efficiency throughout these.
By now, a large variety of open-source OCR tools are available \cite{nazeemOpenSourceOCRLibraries2024}

NLP encompasses a wide range of techniques for processing and understanding human language, that allow to turn unstructured text into structured data that can be more easily processed and retrieved.
While LLMs currently represent the state-of-the-art in NLP, they are not the only option for improving document retrieval, and they may not be the most suitable for our use case, due to their high computational requirements and potential privacy concerns.
Other techniques, such as BM25 and TF-IDF, can be used to improve keyword-based search, while embedding-based approaches can be used to enable semantic search capabilities \cite{patilSurveyTextRepresentation2023}.
Of particular interest to our use case, \textcite{patilSurveyTextRepresentation2023} outlines how domain-specific and knowledge-enriched embeddings are particularly useful in capturing semantic and syntactical relationships, by using local contextual information and external knowledge sources, such as Wikipedia.

Search and retrieval systems have also seen significant advancements, with the development of more sophisticated indexing and ranking algorithms, as well as the integration of NLP techniques to enhance search capabilities.
\cite{shalagin2024survey}

Lastly, while metadata has become a standard tool for enriching digital objects with structured information, there is a lack of standardization in the field, which hinders interoperability and requires case-by-case solutions for metadata extraction and structuring \cite{ulrichUnderstandingNatureMetadata2022}.
While this thus suggests that we need to develop a custom solution for our use case, we take inspiration from domain-relevant implementations, such as the work of ...

\subsection{Aim of the Project}


The objective of this project is to design a working model that can enhance the retrieval of engineering documents concerning civil infrastructure assets, such as bridges, viaducts and culverts. 
At the moment, asset specialists are forced to conduct a manual search in a number of different and inconsistent archival directories that use varying naming conventions, classifications, and terminologies. 
This has resulted in a situation where the retrieval of specific construction drawings and technical documents is a slow task. 

This project has been created to address this situation and provide a solution that can make document discovery a quicker and more assured process. 
The project will involve designing a model as a proof of concept that will help in searching documents in a better and more efficient manner. 
The project will make use of artificial intelligence technologies that can process and retrieve documents, making the task simpler and quicker for specialists. 
The main objective is to design a solution that is simple and easy to use for the end users, who already possess the required expertise and knowledge in the domain. 
Furthermore, our team will try to ensure that the proposed solution is sustainable, privacy-preserving and cost-effective. 
Therefore the system will utilize open-source tools, which will ensure the privacy of the information contained within the archives. 

In conclusion, the project aims to provide an intelligent yet lightweight retrieval system that can significantly improve the efficiency of searching engineering archives, therefore supporting better maintenance planning, structural analysis and safety assessments for civil engineering assets. 
By reducing the time required to find critical documentation and minimizing the risk of missing important records, the system promises to improve the work of asset specialists and improve decision-making processes. 
This improvement can lead to more reliable infrastructure management, faster project preparation and reduced operational delays.

\section{Project Plan}

\subsection{Approach}

The project will follow an iterative, sprint-based development approach combining exploratory data analysis, prototype development, and progressive system refinement. The overall strategy is to move from understanding the archival structure and its limitations toward building a proof of concept that enhances document retrieval through artificial intelligence techniques.

We aim to develop a low-cost, efficient, and user-friendly system, which is aimed at specialist users, familiar with the data and the domain.
Thus, we emphasise simplicity and high-level user interaction, such as domain-specific filters and keywords to narrow down search results, rather than complex interfaces or advanced features that may not be necessary for the target users.

In that vein, we aim to avoid the use of LLMs as well as the use of cloud-based proprietary tools, to ensure sustainability and data sovereignty.
While a RAG-type approach is becoming a standard in search engines, based on ethical and sustainability concerns we strive to find alternatives to such solutions.
We expect that adding this would not improve performance and would thus not be worth the human and ecological costs required for this feature. 
Additionally, we aim to recommend the deployment of the system on a local server, to avoid any privacy concerns related to the potentially sensitive information contained in the documents.
To that end, we will use open-source tools and libraries that can feasibly be run locally, and we will include recommendations for hardware requirements and deployment strategies in our final report, along with a discussion of the trade-offs between local and cloud-based deployment options, and our motivations for choosing a local deployment.




\subsection{Methods}

The methodology of this project combines classical information retrieval techniques with modern natural language processing (NLP) approaches to improve archival document search.

\subsubsection{Document Retrieval}
The project is dependant on document retrieval. Users can use queries to search certain keywords about a project (i.e. regions, place names, specific structure types etc.)
to retrieve documents that are related to these keywords. For this we will use the BM25 formula, which is a well-established equation to score the relevancy of documents.
We will also be creating an index for the documents. An index makes it so certain words are already linked to certain documents, speeding up current retrieval time and making the process more efficient.

A possible extension that we are considering is using semantic search. With semantic search, our process is to create embeddings over both the query and the document and match these two.
With current transformer-based embedding models it has shown to improve performance over BM25, however these models are more energy and time intensive.

\subsubsection{Data extraction}
Before we start with working on the document retrieval, we first need quality textual data. 
Since most of the PDF files that we will be working with are quite old, we will need to somehow extract any type of textual data from these files. 
We will be using Optical Character Recognition (OCR) methods to extract this text.

Besides this, the data from this model will have to be cleaned. This will be dependant on the quality of the model that we use and the amount of processing already in that model.
We will also be looking at retaining any structures within PDF files, such as tables and legends, as this can increase performance for retrieval.
Another possible extension, if time allows, is looking at classifying the documents into specific document types. Examples of this would be: technical drawings, reports, articles etc.
This would then be included in the data to improve retrieval if queries include these terms.



\subsection{Design}

The system will follow a modular architecture consisting of the following components:
\begin{itemize}
    \item Data Ingestion and Preprocessing: handles the ingestion of archival documents, including OCR processing for scanned documents, and preprocessing steps such as text cleaning and normalization
    \item Metadata Extraction and Classification: extracts relevant metadata from the documents (e.g., asset type, construction date, location) and classify documents based on their content and metadata.     
    \item Search Engine: keyword-based and semantic search capabilities, allowing users to retrieve documents based on specific terms as well as conceptual understanding.
    \item User Interaction Interface: simple interface that allows asset specialists to input search queries, apply filters, and view results.
\end{itemize}

As proof of concept we aim to implement a functional prototype of the system that demonstrates the core functionalities, including document ingestion, metadata extraction, and search capabilities, based on a small test dataset of archival documents.

The system performs computational tasks along two different pipelines; The first extracts information from the existing documents followed by indexing and potential embedding into a vector-based database, and will handle the integration of any new documents into the archive.
The other pipeline handles the retrieval of items from the archive, using BM25 and keyword search to retrieve the relevant documents. A potential extension could be also embedding the query with something like BGE-M3 to improve results using semantic search, this will depend on the performance that we measure.

Other than that, we consider creating the front-end design to something similar to a search engine, where there is a search bar and the relevant documents appear below, ranked on relevancy. 
It would be interesting however to highlight parts of text or take text snippets so the user can confirm in a quick glance if the document is potentially relevant or not.

\subsection{Analyses}

The system will be evaluated both quantitatively and qualitatively.
Quantitative Evaluation depends on the availability of labelled data, as well as the possibility to simulate realistic user interactions. To enable this, we will request the company to provide these resources, or to assist us in creating them based on their domain expertise.
We aim to employ the following metrics to determine whether the system reliably retrieves relevant items and does so in a more efficient way than the current manual search process:

\begin{itemize}
    \item Precision and recall of retrieved documents.
    \item F1-score for classification tasks.
    \item Comparison of baseline keyword search versus semantic search performance.
    \item Measurement of average retrieval time.
    \item Reduction in manual search effort (simulated task comparison).
\end{itemize}


Qualitative Evaluation depends on the availability of realistic test queries as well as user feedback. We also request the company to provide these resources, or to assist us in creating them based on their domain expertise.

We aim to get an estimate of the quality of the system performance based on the following:

\begin{itemize}
    \item Relevance assessment of retrieved documents.
    \item User feedback from asset specialists (if available).
    \item Assessment of usability and clarity of results.
\end{itemize}